---
title: "My Plotly Presentation"
author: "Samuel Quiroga"
date: "November 10, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(plotly)
library(DT)
```

## Assignment Two - Developing Data Products

This is Assignment Two, Developing Data Products of the Data Science Specialization in Coursera. This document outlines how to create a simple plot using the `plotly` package in R.

## Data to Visualize

For this assignment we'll use greenhouse gas emissions data from the Greenhouse Gas Reporting Program in the Government of Canada website. 

The Greenhouse Gas Reporting Program (GHGRP) collects information on greenhouse gas (GHG) emissions annually from facilities across Canada. It is a mandatory program for those who meet the requirements. For more information, check out the website at [https://climate-change.canada.ca/facility-emissions]

## The Plot with Plotly

```{r}
# set url of the data
my.url <- "http://data.ec.gc.ca/data/substances/monitor/greenhouse-gas-reporting-program-ghgrp-facility-greenhouse-gas-ghg-data/PDGES-GHGRP-GHGEmissionsGES-2004-Present.csv"
# read data
ghg <- read.csv(my.url)
# get names
names <- colnames(ghg)
# clean column names (remove French version)
my.names <- if_else(is.na(str_extract(names,".*(?=\\.\\.\\.)")),names,str_extract(names,".*(?=\\.\\.\\.)"))
colnames(ghg) <- my.names
# Select rows of interest
my.ghg <- ghg %>% select(Facility.Province.or.Territory,
                         Reference.Year,
                         Total.Emissions..tonnes.CO2e.) %>%
        mutate(Reference.Year = as.Date(Reference.Year,format = "%Y"),
                Total.Emissions..tonnes.CO2e. = 
                       round(if_else(Total.Emissions..tonnes.CO2e. == 0,1,Total.Emissions..tonnes.CO2e.),1)) %>%
        filter(complete.cases(.))

p <- plot_ly(data = my.ghg, 
             x=~Reference.Year,
             y=~Total.Emissions..tonnes.CO2e.,
             type="bar",
             name=~Facility.Province.or.Territory,
             colors = "YlOrRd") %>%
        layout(yaxis=list(title="Total greenhouse gas emissions (tonnes CO2 eq.)"),barmode="stack")

p
```

## The table of values 

```{r}
sum.ghg <- my.ghg %>% 
        group_by(Facility.Province.or.Territory,Reference.Year) %>% 
        summarize(Sum.per.Province = sum(sum(Total.Emissions..tonnes.CO2e.)))
# view table
datatable(sum.ghg,options = list(scrollX=TRUE,pageLength = 5))
```